From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-07-22 dictionary.tex
Deleting lots of comments

Motivated by an application to quasicharacters of algebraic tori over local fields, in this paper we extend the
function-sheaf dictionary from commutative algebraic groups over finite fields to smooth commutative group schemes $G$ over $\Fq$.
In order to do this, we replace the local system $\cs{E}$ on $G$ with a Weil local system while retaining the extra structure $\mu_\cs{E}$. 
In this way we are led to the category $\CS(G)$ of {\it character sheaves} on $G$ (\S\ref{ssec:category}):
objects in $\CS(G)$ are triples $(\gcs{L}, \mu,\phi)$, where $\gcs{L}$ is a rank-one local system on $\bG := G\times_{\Spec{\Fq}} \Spec{\bFq}$
and $\phi : \Frob{G}^* \gcs{L}\to \gcs{L}$ and $\mu : \bm^* \gcs{L} \iso \gcs{L} \boxtimes \gcs{L}$ are
isomorphisms of sheaves satisfying certain compatibility conditions;
morphisms in $\CS(G)$ are then morphisms of Weil sheaves which are compatible with the extra structure.
This paper establishes the basic properties of category $\CS(G)$, using the group homomorphism 
\[
\TrFrob{G}: \CSiso{G} \to G(\Fq)^*
\]
provided by the trace of Frobenius  to find the relation between character sheaves on $G$ and characters of $G(\Fq)$.
Then we return to our motivating application and use character sheaves to geometrize and categorify quasicharacters algebraic tori over local fields.
% (\S \ref{sec:applications}).

@@

Next, we consider character sheaves on \'etale commutative group schemes $G$ over $\Fq$ (\S\ref{sec:etale}).
\'Etale group schemes form a counterpoint to connected algebraic groups, since the component group of any
smooth group scheme is an \'etale group scheme.
Our key tools for understanding the trace of Frobenius in the \'etale case 
are a reinterpretation of $\CS(G)$ in terms of stalks (\S\ref{ssec:stalks}) and 
the Hochschild-Serre spectral sequence (\S\ref{ssec:E}) for $\Weil{} \ltimes \bG$, where $\Weil{} \subset \Gal(\Fq/\Fq)$ is the Weil group for $\Fq$.
We define (\S\ref{ssec:S}) an isomorphism $S_G$ from $\CSiso{G}$ to the second cohomology of the total space of the spectral sequence
\[
E_2^{p,q} \ceq \Hh^p(\Weil{}, \Hh^q(\bG, \EEx)) \Rightarrow \Hh^{p+q}(\Weil{} \ltimes \bG, \EEx).
\]
Paired with the short exact sequence
\[
  0 \to \Hh^0(\Weil{},\Hh^2(\bG,\EEx)) \to \Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1(\bG,\EEx)) \to 0
\]
arising from the spectral sequence, the isomorphism $S_G$ allows us to show (\S\ref{ssec:SandT})
that the group homomorphism $\TrFrob{G}: \CSiso{G} \to G(\Fq)^*$ is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$.
The necessity of using Weil local systems on $G$ in the definition of $\CS(G)$ already appears here:
if one were to use local systems on $G$ instead, the group homomorphism $\TrFrob{G}$ would not then be surjective (\S \ref{ssec:bS}). 
Moreover, as examples show (\S \ref{eg:}), the kernel of $\TrFrob{G}$ is non-trivial in general.
%The fact that $\TrFrob{G}$ may not be an isomorphism should not be too troubling:
%if quasicharacter sheaves $(\gcs{L}, \mu,\phi)$ and $(\gcs{L}', \mu',\phi')$ determine the same character
%of $G(\Fq)$ then $(\gcs{L},\phi)$ and $(\gcs{L}',\phi')$ are isomorphic as Weil sheaves.

%The appearance of $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ in these two cases is interesting: on the one hand,
%when $G$ is a connected commutative algebraic group over $\Fq$, isomorphism classes of quasicharacter
%sheaves on $G$ are measured by $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ and the trace of Frobenius gives an
%isomorphism $\TrFrob{G}: \CSiso{G} \to \Hom(G(\Fq),\EEx)$; on the other hand, if $G$ an \'etale commutative
%group scheme over $\Fq$, then $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ measures the kernel of $\TrFrob{G}: \CSiso{G} \to \Hom(G(\Fq),\EEx)$. 
%This observation makes plain the fact that  . . .
%Indeed, quasicharacter sheaves on \'etale commutative group schemes that are trivialized
%by discrete isogenies are invisible to the trace of Frobenius.

Having understood $\CS(G)$ in two extreme cases --
for connected commutative algebraic groups and for \'etale commutative group schemes --
we turn to the case of smooth commutative group schemes (\S\ref{sec:main}) using the component group sequence
\[
0 \to G^0 \to G \to \pi_0(G) \to 0.
\]
Using pullbacks of character sheaves we obtain a diagram
\[
  \begin{tikzcd}[row sep=20, column sep=20]
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar & G(\Fq)^* \rar & G^0(\Fq)^* \rar & 0.
  \end{tikzcd}
\]
We show that the rows of this diagram are exact (\S\S\ref{ssec:restriction},\ref{ssec:component}),
so we may apply the snake lemma to prove the main theorem of the paper,
\begin{theorem*}[{Thm. \ref{thm:snake}}]
If $G$ is a smooth commutative group scheme over $\Fq$ then the trace of Frobenius gives a short exact sequence
\[
\begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \CSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0.
\end{tikzcd}
\]
\end{theorem*}
\noindent
Thus, unless the component group scheme $\pi_0(\bG)$ is cyclic, category $\CS(G)$ contains
non-trivial character sheaves lying in the kernel of the trace of Frobenius and therefore lost in translation to characters of $G(\Fq)$.
%These {\it invisible character sheaves} are parametrized by $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$.

@@

Finally, specializing to the case that $X = T$ is the N\'eron model of an algebraic torus over $K$ (\S\ref{ssec:CS_tori}), 
we give a canonical short exact sequence 
\[
0 \to \Hh^2(X_*(T)_{\mathcal{I}_K},\EEx)^{\Weil{}} \to \QCSiso{T} \to \Hom_\text{}(T(K),\EEx) \to 0,
\]
where $X_*(T)_{\mathcal{I}_K}$ is the group of coinvariants of the cocharacter lattice $X_*(T)$ of the algebraic torus
$T_K$ by the action of the inertia group $\mathcal{I}_K$ of $K$, and where $\Hom_\text{}(T(K),\EEx)$
denotes the group of quasicharacters of $T(K)$.
We further show that automorphism groups in $\QCS(T)$ are given, for every quasicharacter sheaf $\cs{F}$ for $T$, by
\[
\Aut(\cs{F}) \iso (\check{T}_\ell)^{\Weil{K}},
\]
where $\Weil{K}$ is the Weil group for $K$ and $\check{T}_\ell$ is the $\ell$-adic dual torus to $T$.
%The appearance of the dual torus here is quite striking.

@@

%We also want to situate our terminology in a historical context.
For a connected commutative algebraic group over a finite field, it is not uncommon to refer to
local systems satisfying \eqref{introbox}  as character sheaves; see for example, \cite{kamgarpour:09a}*{Intro}.
Our definition of character sheaves on smooth commutative group schemes over finite fields evolved from this notion,
with an eye toward quasicharacters of $p$-adic groups.
The process of creating a category from the group of quasicharacters of a $p$-adic torus informs our choice of the
term quasicharacter sheaf in this paper.

@@

We will make use of the short exact sequence of smooth group schemes defining the component group scheme for $G$:
\[
\begin{tikzcd}
0 \rar & G^0 \arrow{r}{\iota_0} & G \arrow{r}{\pi_0} & \pi_0(G) \rar & 0.
\end{tikzcd}
\]
Then $G^0$ is a connected algebraic group and $\pi_0(G)$ is an \'etale commutative group scheme.
In contrast to the case of algebraic varieties, the component group scheme $\pi_0(G)$ for $G$ need not be finite.
%Unless noted otherwise, we place absolutely no conditions on the \'etale group scheme $\pi_0(G)$.

@@

\begin{definition}\label{def:CS}
A \emph{character sheaf on $G$} is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.1)}{CS.1} $\gcs{L}$ is a rank-one $\ell$-adic local system on $\bG$, by which we mean a constructible
$\ell$-adic sheaf on $\bG$, {\it lisse} on each connected component of $\bG$, whose stalks are one-dimensional $\EE$-vector spaces;
\labitem{(CS.2)}{CS.2} $\mu: \bm^* \gcs{L} \to \LxL$ is an isomorphism of
sheaves on $\bGxG$ such that the following diagram commutes,
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL
  \end{tikzcd}
  \]
\labitem{(CS.3)}{CS.3} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with
  $\mu$ in the sense that the following diagram commutes.
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{\phi\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL
  \end{tikzcd}
  \]
\end{enumerate}
%\end{definition}
Morphisms of character sheaves are defined in the natural way:
\begin{enumerate}
\labitem{(CS.4)}{CS.4} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are character sheaves on $G$ then
  a morphism $\rho : \cs{L} \to \cs{L}'$ is a map $\brho : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \brho} \arrow[swap]{d}{\phi} & \Frob{G}^* \gcs{L'} \dar{\phi'}
  & & \arrow[swap]{d}{\mu} \bm^* \gcs{L} \rar{\bm^* \brho} & \bm^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\brho} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\rho\boxtimes \rho}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}
 The category of character sheaves on $G$ will be denoted by $\CS(G)$.
 \end{definition}

@@

We will describe the group $\CSiso{G}$ in Theorem~\ref{thm:snake}
and the sets $\Hom(\cs{L},\cs{L}')$ in Theorem~\ref{thm:autornaught}; in this way we provide a complete description of the category $\CS(G)$.
%
In the meantime, we make an elementary observation about $\Hom(\cs{L},\cs{L}')$.

@@

The next two results follow easily from the definitions.
%In the remainder of this section We devote the remainder of this section to studying the behavior
%of the trace of Frobenius under morphisms of group schemes.  These simple but important properties of
%pullbacks and external products will be used in the proofs of many results in this paper.

@@

\iffalse
\begin{proof}
The only non-trivial part is the last claim, so we will only address that point here.
%
Set $G \ceq G_1\times G_2$
and write $e_1$ and $e_2$ for the identity elements of $G_1$ and $G_2$.
Define $f : G\to G\times G$ by $f(g_1,g_2) \ceq (g_1,e_2,e_1,g_2)$.
Observe that $m\circ f = \id_G$.
Let $p_1$, $p_2$ be the projection morphisms pictured below:
\[
\begin{tikzcd}
G & \arrow[swap]{l}{p_1} G\times G \arrow{r}{p_2} & G.
\end{tikzcd}
\]
Let $r_1$ and $r_2$ be the projection morphisms pictured below,
with sections $q_1$ and $q_2$, also morphisms of group schemes:
\[
\begin{tikzcd}
G_1  \arrow[swap, bend right]{r}{q_1} &
\arrow[swap, bend right]{l}{r_1} G_1\times G_2 \arrow[bend left]{r}{r_2} &
\arrow[bend left]{l}{q_2} G_2.
\end{tikzcd}
\]
Observe that $p_1\circ f = q_1 \circ r_1$ and $p_2 \circ f = q_2\circ r_2$.
%
Now, let $\cs{L} \ceq (\gcs{L},\mu,\phi)$ be a character sheaf on $G$
and set $\cs{L}_1 \ceq q_1^* \cs{L}$ and $\cs{L}_2 \ceq q_2^* \cs{L}$.
By Lemma~\ref{lem:pullback}, $\cs{L}_1$ is a character sheaf on $G_1$
and $\cs{L}_2$ is a character sheaf on $G_2$.
We will obtain an isomorphism $\cs{L} \iso  \cs{L}_1\boxtimes \cs{L}_2$.

@@

%\subsection{character sheaves on connected commutative algebraic groups} \label{sec:disc-isog}

@@

%Before turning our attention to character sheaves on connected commutative algebraic groups in Section~\ref{ssec:connected}, 
%Here we treat two special classes of character sheaves on arbitrary smooth commutative group schemes $G$:
%those that descend from local systems on $\bG$ to local systems on $G$ (\S \ref{ssec:descentG});
%and those that are defined by discrete isogenies onto $G$ (\S \ref{ssec:discrete_isogenies}).
%The latter will play a role in Sections~\ref{ssec:connected} and \ref{ssec:restriction}.

@@

%\subsection{The function-sheaf dictionary for connected algebraic groups over finite fields}\label{ssec:connected}

@@

\iffalse
\begin{proof}
Choose any $\Fq$-rational point $g$ on $G$ and let $\bg$ be the geometric point on $G$ lying above $g$.
Recall that the \emph{Weil group} of $G$, which we will denote by $\Weil{}(G,\bg)$, is a subgroup of the \'etale
fundamental group defined by the following diagram:
\[
 \begin{tikzcd}
 1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \Weil{}(G,\bg) \rar \dar[hook] & \Weil{} \rar \dar[hook] & 1 \\
 1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \rar & 1.
 \end{tikzcd}
\]
The $\Fq$-rational point $g$ under the geometric point $\bg$ determines a splitting
$\Weil{}\to \Weil{}(G,\bg)$ of $\Weil{}(G,\bg)\to \Weil{}$.
%
  Since $G$ is connected, the geometric point $\bg$ determines
  an equivalence between the category of $\ell$-adic Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil{}(G,\bg)$ \cite{deligne:80a}*{1.1.12}.

  Now let $(\gcs{L},\mu,\phi)$ be a character sheaf on $G$
  and let $\lambda : \Weil{}(G, \bg) \to \EEx$ be the character determined by $(\gcs{L},\phi)$.
  Composing with the splitting $\Weil{} \to \Weil{}(G,\bg)$ yields an $\ell$-adic character
  $\lambda_g : \Weil{} \to \EEx$, which is the same as the Trace of Frobenius defined in Section~\ref{ssec:Frob}, for every $\Fq$ rational point $g$ on $G$:
  $
  \lambda_g(\Frob{}) =  \trFrob{\cs{L}}(g).
  $
%
  On the other hand, we have already seen that $\trFrob{\cs{L}} : G(\Fq) \to \EEx$
  is a group homomorphism. 
  Since $G$ is an algebraic group over $\Fq$, $G(\Fq)$ is finite.
  Therefore $\trFrob{\cs{L}}(g) = \lambda_g(\Frob{})$ is a root of unity
  for every $g\in G(\Fq)$.  Since $\Weil{}$ is generated by
  $\Frob{}$ and $\lambda_g : \Weil{} \to \EEx$ is
  a character, it follows that the image of $\lambda_g$ is a finite group.
  Thus, $\lambda_g$ extends to an $\ell$-adic character of $\Gal(\bFq/\Fq)$,
  which we will also denote $\lambda_g$.
%
  We may now lift the $\ell$-adic character $\lambda_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$. 
 % 
  The $\Fq$ rational point $g$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\cs{E}$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\cs{E} \iso \gcs{L}$.

@@

The equivalence $G \mapsto G(\bFq)$,
from the category of \'etale commutative group schemes over $\Fq$ to the category of commutative groups equipped
with a continuous action of $\Gal(\bFq/\Fq)$,
provides the following simple description of character sheaves.
%
A character sheaf $\cs{L}$ on an \'etale commutative group scheme $G$ over $\Fq$ is:
\begin{enumerate}
 \labitem{(cs.1)}{cs.1} an indexed set of one-dimensional
  $\EE$-vector spaces $\gcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

>>>>

Motivated by an application to quasicharacters of algebraic tori over local fields, in this paper we extend the
function-sheaf dictionary from commutative algebraic groups over finite fields to smooth commutative group schemes $G$ over $\Fq$.
In order to do this, we replace the local system $\cs{E}$ on $G$ with a Weil local system while retaining the extra structure $\mu_\cs{E}$. 
In this way we are led to the category $\CS(G)$ of {\it character sheaves} on $G$ (\S\ref{ssec:category}):
objects in $\CS(G)$ are triples $(\gcs{L}, \mu,\phi)$, where $\gcs{L}$ is a rank-one local system on $\bG := G\times_{\Spec{\Fq}} \Spec{\bFq}$
and $\phi : \Frob{G}^* \gcs{L}\to \gcs{L}$ and $\mu : \bm^* \gcs{L} \iso \gcs{L} \boxtimes \gcs{L}$ are
isomorphisms of sheaves satisfying certain compatibility conditions;
morphisms in $\CS(G)$ are then morphisms of Weil sheaves which are compatible with the extra structure.
This paper establishes the basic properties of category $\CS(G)$, using the group homomorphism 
\[
\TrFrob{G}: \CSiso{G} \to G(\Fq)^*
\]
provided by the trace of Frobenius  to find the relation between character sheaves on $G$ and characters of $G(\Fq)$.
Then we return to our motivating application and use character sheaves to geometrize and categorify quasicharacters algebraic tori over local fields.

@@

Next, we consider character sheaves on \'etale commutative group schemes $G$ over $\Fq$ (\S\ref{sec:etale}).
\'Etale group schemes form a counterpoint to connected algebraic groups, since the component group of any
smooth group scheme is an \'etale group scheme.
Our key tools for understanding the trace of Frobenius in the \'etale case 
are a reinterpretation of $\CS(G)$ in terms of stalks (\S\ref{ssec:stalks}) and 
the Hochschild-Serre spectral sequence (\S\ref{ssec:E}) for $\Weil{} \ltimes \bG$, where $\Weil{} \subset \Gal(\Fq/\Fq)$ is the Weil group for $\Fq$.
We define (\S\ref{ssec:S}) an isomorphism $S_G$ from $\CSiso{G}$ to the second cohomology of the total space of the spectral sequence
\[
E_2^{p,q} \ceq \Hh^p(\Weil{}, \Hh^q(\bG, \EEx)) \Rightarrow \Hh^{p+q}(\Weil{} \ltimes \bG, \EEx).
\]
Paired with the short exact sequence
\[
  0 \to \Hh^0(\Weil{},\Hh^2(\bG,\EEx)) \to \Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1(\bG,\EEx)) \to 0
\]
arising from the spectral sequence, the isomorphism $S_G$ allows us to show (\S\ref{ssec:SandT})
that the group homomorphism $\TrFrob{G}: \CSiso{G} \to G(\Fq)^*$ is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$.
The necessity of using Weil local systems on $G$ in the definition of $\CS(G)$ already appears here:
if one were to use local systems on $G$ instead, the group homomorphism $\TrFrob{G}$ would not then be surjective (\S \ref{ssec:bS}). 
Moreover, as examples show (\S \ref{eg:}), the kernel of $\TrFrob{G}$ is non-trivial in general.

Having understood $\CS(G)$ in two extreme cases --
for connected commutative algebraic groups and for \'etale commutative group schemes --
we turn to the case of smooth commutative group schemes (\S\ref{sec:main}) using the component group sequence
\[
0 \to G^0 \to G \to \pi_0(G) \to 0.
\]
Using pullbacks of character sheaves we obtain a diagram
\[
  \begin{tikzcd}[row sep=20, column sep=20]
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar & G(\Fq)^* \rar & G^0(\Fq)^* \rar & 0.
  \end{tikzcd}
\]
We show that the rows of this diagram are exact (\S\S\ref{ssec:restriction},\ref{ssec:component}),
so we may apply the snake lemma to prove the main theorem of the paper,
\begin{theorem*}[{Thm. \ref{thm:snake}}]
If $G$ is a smooth commutative group scheme over $\Fq$ then the trace of Frobenius gives a short exact sequence
\[
\begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \CSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0.
\end{tikzcd}
\]
\end{theorem*}
If the component group scheme $\pi_0(\bG)$ is cyclic, then the kernel of $\TrFrob{G}$ will be trivial and
each character of $G(\Fq)$ will uniquely determine a character sheaf on $G$, up to isomorphism.  But when
$\pi_0(\bG)$ is large (c.f. Remark \ref{rem:H2}), $G$ will admit \emph{invisible character sheaves} with
trivial trace of Frobenius.

@@

Finally, specializing to the case that $X = T$ is the N\'eron model of an algebraic torus over $K$ (\S\ref{ssec:CS_tori}), 
we give a canonical short exact sequence 
\[
0 \to \Hh^2(X_*(T)_{\mathcal{I}_K},\EEx)^{\Weil{}} \to \QCSiso{T} \to \Hom_\text{}(T(K),\EEx) \to 0,
\]
where $X_*(T)_{\mathcal{I}_K}$ is the group of coinvariants of the cocharacter lattice $X_*(T)$ of the algebraic torus
$T_K$ by the action of the inertia group $\mathcal{I}_K$ of $K$, and where $\Hom_\text{}(T(K),\EEx)$
denotes the group of quasicharacters of $T(K)$.
We further show that automorphism groups in $\QCS(T)$ are given, for every quasicharacter sheaf $\cs{F}$ for $T$, by
\[
\Aut(\cs{F}) \iso (\check{T}_\ell)^{\Weil{K}},
\]
where $\Weil{K}$ is the Weil group for $K$ and $\check{T}_\ell$ is the $\ell$-adic dual torus to $T$.

@@

For a connected commutative algebraic group over a finite field, it is not uncommon to refer to
local systems satisfying \eqref{introbox}  as character sheaves; see for example, \cite{kamgarpour:09a}*{Intro}.
Our definition of character sheaves on smooth commutative group schemes over finite fields evolved from this notion,
with an eye toward quasicharacters of $p$-adic groups.
The process of creating a category from the group of quasicharacters of a $p$-adic torus informs our choice of the
term quasicharacter sheaf in this paper.

@@

We will make use of the short exact sequence of smooth group schemes defining the component group scheme for $G$:
\[
\begin{tikzcd}
0 \rar & G^0 \arrow{r}{\iota_0} & G \arrow{r}{\pi_0} & \pi_0(G) \rar & 0.
\end{tikzcd}
\]
Then $G^0$ is a connected algebraic group and $\pi_0(G)$ is an \'etale commutative group scheme.
In contrast to the case of algebraic varieties, the component group scheme $\pi_0(G)$ for $G$ need not be finite.

@@

\begin{definition}\label{def:CS}
A \emph{character sheaf on $G$} is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.1)}{CS.1} $\gcs{L}$ is a rank-one $\ell$-adic local system on $\bG$, by which we mean a constructible
$\ell$-adic sheaf on $\bG$, {\it lisse} on each connected component of $\bG$, whose stalks are one-dimensional $\EE$-vector spaces;
\labitem{(CS.2)}{CS.2} $\mu: \bm^* \gcs{L} \to \LxL$ is an isomorphism of
sheaves on $\bGxG$ such that the following diagram commutes,
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL
  \end{tikzcd}
  \]
\labitem{(CS.3)}{CS.3} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with
  $\mu$ in the sense that the following diagram commutes.
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{\phi\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms of character sheaves are defined in the natural way:
\begin{enumerate}
\labitem{(CS.4)}{CS.4} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are character sheaves on $G$ then
  a morphism $\rho : \cs{L} \to \cs{L}'$ is a map $\brho : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \brho} \arrow[swap]{d}{\phi} & \Frob{G}^* \gcs{L'} \dar{\phi'}
  & & \arrow[swap]{d}{\mu} \bm^* \gcs{L} \rar{\bm^* \brho} & \bm^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\brho} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\rho\boxtimes \rho}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}
 The category of character sheaves on $G$ will be denoted by $\CS(G)$.
 \end{definition}

@@

We will describe the group $\CSiso{G}$ in Theorem~\ref{thm:snake}
and the sets $\Hom(\cs{L},\cs{L}')$ in Theorem~\ref{thm:autornaught}; in this way we provide a complete description of the category $\CS(G)$.
In the meantime, we make an elementary observation about $\Hom(\cs{L},\cs{L}')$.

@@

The next two results follow easily from the definitions.

@@

\iffalse
\begin{proof}
The only non-trivial part is the last claim, so we will only address that point here.
Set $G \ceq G_1\times G_2$
and write $e_1$ and $e_2$ for the identity elements of $G_1$ and $G_2$.
Define $f : G\to G\times G$ by $f(g_1,g_2) \ceq (g_1,e_2,e_1,g_2)$.
Observe that $m\circ f = \id_G$.
Let $p_1$, $p_2$ be the projection morphisms pictured below:
\[
\begin{tikzcd}
G & \arrow[swap]{l}{p_1} G\times G \arrow{r}{p_2} & G.
\end{tikzcd}
\]
Let $r_1$ and $r_2$ be the projection morphisms pictured below,
with sections $q_1$ and $q_2$, also morphisms of group schemes:
\[
\begin{tikzcd}
G_1  \arrow[swap, bend right]{r}{q_1} &
\arrow[swap, bend right]{l}{r_1} G_1\times G_2 \arrow[bend left]{r}{r_2} &
\arrow[bend left]{l}{q_2} G_2.
\end{tikzcd}
\]
Observe that $p_1\circ f = q_1 \circ r_1$ and $p_2 \circ f = q_2\circ r_2$.
Now, let $\cs{L} \ceq (\gcs{L},\mu,\phi)$ be a character sheaf on $G$
and set $\cs{L}_1 \ceq q_1^* \cs{L}$ and $\cs{L}_2 \ceq q_2^* \cs{L}$.
By Lemma~\ref{lem:pullback}, $\cs{L}_1$ is a character sheaf on $G_1$
and $\cs{L}_2$ is a character sheaf on $G_2$.
We will obtain an isomorphism $\cs{L} \iso  \cs{L}_1\boxtimes \cs{L}_2$.

@@

@@

@@

@@

\iffalse
\begin{proof}
Choose any $\Fq$-rational point $g$ on $G$ and let $\bg$ be the geometric point on $G$ lying above $g$.
Recall that the \emph{Weil group} of $G$, which we will denote by $\Weil{}(G,\bg)$, is a subgroup of the \'etale
fundamental group defined by the following diagram:
\[
 \begin{tikzcd}
 1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \Weil{}(G,\bg) \rar \dar[hook] & \Weil{} \rar \dar[hook] & 1 \\
 1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \rar & 1.
 \end{tikzcd}
\]
The $\Fq$-rational point $g$ under the geometric point $\bg$ determines a splitting
$\Weil{}\to \Weil{}(G,\bg)$ of $\Weil{}(G,\bg)\to \Weil{}$.
  Since $G$ is connected, the geometric point $\bg$ determines
  an equivalence between the category of $\ell$-adic Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil{}(G,\bg)$ \cite{deligne:80a}*{1.1.12}.

  Now let $(\gcs{L},\mu,\phi)$ be a character sheaf on $G$
  and let $\lambda : \Weil{}(G, \bg) \to \EEx$ be the character determined by $(\gcs{L},\phi)$.
  Composing with the splitting $\Weil{} \to \Weil{}(G,\bg)$ yields an $\ell$-adic character
  $\lambda_g : \Weil{} \to \EEx$, which is the same as the Trace of Frobenius defined in Section~\ref{ssec:Frob}, for every $\Fq$ rational point $g$ on $G$:
  $
  \lambda_g(\Frob{}) =  \trFrob{\cs{L}}(g).
  $
  On the other hand, we have already seen that $\trFrob{\cs{L}} : G(\Fq) \to \EEx$
  is a group homomorphism. 
  Since $G$ is an algebraic group over $\Fq$, $G(\Fq)$ is finite.
  Therefore $\trFrob{\cs{L}}(g) = \lambda_g(\Frob{})$ is a root of unity
  for every $g\in G(\Fq)$.  Since $\Weil{}$ is generated by
  $\Frob{}$ and $\lambda_g : \Weil{} \to \EEx$ is
  a character, it follows that the image of $\lambda_g$ is a finite group.
  Thus, $\lambda_g$ extends to an $\ell$-adic character of $\Gal(\bFq/\Fq)$,
  which we will also denote $\lambda_g$.
  We may now lift the $\ell$-adic character $\lambda_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$. 
  The $\Fq$ rational point $g$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\cs{E}$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\cs{E} \iso \gcs{L}$.

@@

The equivalence $G \mapsto G(\bFq)$,
from the category of \'etale commutative group schemes over $\Fq$ to the category of commutative groups equipped
with a continuous action of $\Gal(\bFq/\Fq)$,
provides the following simple description of character sheaves.
A character sheaf $\cs{L}$ on an \'etale commutative group scheme $G$ over $\Fq$ is:
\begin{enumerate}
 \labitem{(cs.1)}{cs.1} an indexed set of one-dimensional
  $\EE$-vector spaces $\gcs{L}_x$, as $x$ runs over
  $G(\bFq)$;
